---
layout: exercise
title: Basic R 1 with list
---

The number of birds banded at a series of sampling sites has been counted by
your field crew and entered into the following list. The first item in each
sublist is an alphanumeric code for the site and the second value is the number
of birds banded. Cut and paste the list into your assignment and then answer the
following questions by printing them to the screen.

```{r}
    data <- list(c('A1', 28), c('A2', 32), c('A3', 1), c('A4', 0),
                  c('A5', 10), c('A6', 22), c('A7', 30), c('A8', 19),
			      c('B1', 145), c('B2', 27), c('B3', 36), c('B4', 25),
			      c('B5', 9), c('B6', 38), c('B7', 21), c('B8', 12),
			      c('C1', 122), c('C2', 87), c('C3', 36), c('C4', 3),
			      c('D1', 0), c('D2', 5), c('D3', 55), c('D4', 62),
			      c('D5', 98), c('D6', 32))
```

1.  How many sites are there?
```{r echo=FALSE}
  paste('There are ', length(data), ' sites.')
```

2.  How many birds were counted at the 7th site?
```{r echo=FALSE}
  paste(data[[7]][2], ' birds were counted at the 7th site.')
```

3.  How many birds were counted at the last site?
```{r echo=FALSE}
  paste(data[[length(data)]][2], ' birds were counted at the last site.')
```

4.  What is the total number of birds counted across all sites?
```{r echo=FALSE}
  bird_counts <- vector( length = length(data))
  for (i in 1:length(data)){ bird_counts[i] = as.numeric(data[[i]][2]) }

  paste('A total of ', sum(bird_counts), ' birds were counted across all sites.')
```

5.  What is the average number of birds seen on a site?
```{r echo=FALSE}
  paste('The average number of birds is ', round(mean(bird_counts), 0), ' per site.')
```

6.  What is the total number of birds counted on sites with codes
    beginning with C? (don't just identify this sites by eye, in the
    real world there could be hundreds or thousands of sites)

```{r echo=FALSE}
  C_counts = 0
  for (i in 1:length(data)){
    if (substr(data[[i]][1], 1, 1) == "C"){
      C_counts = C_counts + as.numeric(data[[i]][2])}
  }

  paste('A total of ', C_counts, ' birds were counted at sites with codes beginning with C.')
```
