---
title: "Portal population dynamics"
author: "Instructor's name"
date: "December 1, 2015"
output: html_document
---

Explore patterns in population dynamics at Portal.

## Required Libraries 
```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(scales)
```

## Data

Data is from the [Portal project teaching database](http://figshare.com/articles/Portal_Project_Teaching_Database/1314459) published on Figshare. We need the surveys table for our analysis:

```{r, cache=TRUE}
# Setting cache=TRUE lets you re-use the results of this code chunk in other code
# chunks instead of re-calculating or re-downloading it each time.
download.file("https://ndownloader.figshare.com/files/2292172", "surveys.csv")
data <- read.csv("surveys.csv")
```

## Analyze population time-series

Get the time-series of counts for all species.

```{r}
time_series <-
  data %>%
  group_by(species_id, year) %>%
  summarize(count = n()) %>%
  na.omit()

head(time_series)
```

## Plot the time-series.

```{r, message=FALSE, echo=FALSE, cache=TRUE}
ggplot(time_series, aes(x = year, y = count)) +
  geom_point() +
  geom_line() +
  facet_wrap(~species_id) +
  scale_x_continuous(breaks=pretty_breaks(n=2))
```


## A simple model

```{r, echo=FALSE}
model <- data %>% group_by(year) %>% 
  summarize(count = n()) %>% 
  lm(count ~ year, data = .)

results <- anova(model)
```

We found a marginally significant linear relationship between the total count 
and year (p = `r round(results[["Pr(>F)"]][1], 3)`; see Table 1 for more deails)

```{r, echo=FALSE}
# knitr's `kable` function is a handy way to make nice-looking tables 
# from data frames
knitr::kable(results, caption = "A pretty table")
```

